NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE, PG and PS lipids

### 2017-07-27 Manuscript/manuscript.tex
Add POPG data with Calcium, reported by Jesper Madsen

\todo{Simulation data for systems with negatively charged lipids and CaCl$_2$ to be collected}

>>>>

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge \todo{Should we check this against cationic surfactant experiments \cite{scherer89}?}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPG.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure}

Also dependence of $\beta$-carbon of PG on CaCl$_2$ concentration is compared with
experiments \cite{borle85} in Fig. \ref{PSPGchangesWITHCaCl}. Absolute value of
the order parameter is too large without ions, but rapid decrease due to addition of
CaCl$_2$ is observed in agreement with experiments for systems with 1:1 mixture of POPC and POPG.
In addition, absolute value in systems with CaCl$_2$ is in agreement with experiments.
However, system with 4:1 mixture of POPC and POPG behaves differently, but experimental
data is not available for comparison for this mixture.

\todo{More simulation data for systems with negatively charged lipids and CaCl$_2$ to be collected}
